---
title: "A short-Cut MSE-(lite) approach for Harvest Control Rule evaluation towards quotas from annual benchmark assessment advice  for blackspot seabream"

author: 
 - GFCM WGMSE-SBR2 
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    number_sections: yes
    toc: yes
    #keep_tex: false
vignette: >
  %\VignetteIndexEntry{ss3daigs Vignetted}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r wrap-hook, echo = FALSE,warning=FALSE,message=FALSE}
library(knitr)
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})
```


```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "  " ,fig.align = 'center', cache=FALSE,tidy.opts=list(width.cutoff=55), tidy=TRUE)

```




Load packages


```{r, message=FALSE,warning=FALSE}
# Load
library(ggplot2)
library(FLCore)
library(ggplotFL)
library(mse)
library(FLRef)
library(ggpubr)
library(mseviz)

```


# A short-Cut MSE-"lite" approach for Harvest Control Rule (HCR) evaluation with feedback control of emulated annual benchmark advice

In contrast to a full Management Strategy Evaluation (MSE) simulation design (Punt et al. 2017), the MSE  ‘shortcut’ approach, omits the step of the annual updating of the estimation model (assessment) in the feedback control.  Instead, it mimics an annual update of the benchmark assessment model by passing outcomes (SSB and F) from the ‘true’ age-structured dynamics from the operating model (OM) with assessment error to the harvest control rule (HCR) and catch implementation system.


The HCRs were implemented using a simulated feedback control loop between the implementation system and the operating model, where the implementation system translates the emulated assessment outcome via the HRC into  the catch advice. The feedback control loop between the implementation system and the OM allows accounting for the lag between the last of year data used in the assessment and the implementation year of catch advice ($C_{adv}$). 

For blackspot seabream, the implementation system of the harvest control rule is based on the assumption that advice is given for year $y+1$ based on an assessment completed in year $y$, which is fitted to data up until last data year $y-1$. Therefore implementation of the  derived $C_{adv}$ through HCR requires projection of the stock dynamics by way of a short-term forecast. To do this, numbers-at-age were projected through the year of assessment. Status quo recruitment, $M_a$, $w_a$ and $mat_a$ were set as the mean of the last 3 years. A projection based on a fixed fishing mortality-at-age to the last year ($y-1$) in the assessment is then made through to the implementation year ($y+1$).

The limitations of the MSE short-cut approach are that it cannot fully account for uncertainties resulting from imperfect sampling of the full age-structure (e.g. poorly sampled recruits), observation error, misspecified model assumptions. However, these impacts can be to some extend accounted by simulating assessment errors ($SSB$ and $F$), including robustenss tests to account for systematic bias and also for environmental driven processes. The advantage of the short-cut MSE-"lite" approach is straight-forward to implement (`FLR` or `openMSE`) and reduced complexity and computation time when the focus is predominantly optimizing HCRs for setting quotas on the premises that a benchmark assessment form the basis for the advice.

## Objectives

The integrated assessment of blackspot seabream in the Strait of Gibraltar has been benchmarked following the finalization of a rigorous benchmark process. The assessment forms the basis for advice and is consistent with the perception about stock from all stakeholder groups. Furthermore, the benchmark is also informed by additional size information and represent the “base case” of current information about the stock.

A key priority is to transition from the interim management measure (catch and limits limits) to the long-term management plan through 2030. For this purpose,  a management strategy to provide annual catch advice with feedback control (assessment updates + harvest control rule). Considering this, it may be deemed most feasible to take the existent benchmark model as the OM, mimic its updates with assessment error and predominantly focus on a robust harvest control rule (HCR) that enables to compute future catch advice from the updated benchmark in line with the Multiannual Management Plan. This outcome-oriented MSE "lite" approach can facilitate the finalization of the MSE process within the time frame to the roadmap. 

## Glossary

The following glossary summarizes key HCR parameters and associated target and limit reference points
that are considered for tuning the candidate HCRs to optimise the trade-offs between maximising fishing
opportunity and risk:


+ $F_{tgt}$: target reference point for fishing mortality at Fmsy (or its proxy), here $F_{B35}$

+ $B_{tgt}$: the average biomass around which the biomass fluctuated when fishing at Ftgt, here $B_{35}$

+ $B_{lim}$: a deterministic biomass limit reference point below which a stock is considered to have reduced reproductive capacity. Here $B_{lim}$ was set to $0.25B_{tgt}$ 

+ $B_{pa}$: a precautionary biomass reference point set with high probability that biomass is above $B_{lim}$, which acts as a safety margin below which the risk of reduced reproductive capacity is increasing. When the biomass is estimated to be above Bpa, the stock is considered to be within safe biological limits in terms of its reproductive capacity. Here $B_{lim}$ was set to $0.5B_{tgt}$ 

+ $F_{adv}$: advised fishing mortality from the HCR, which can be specified as ratio of $F_{tgt}$.

+ $C_{adv}$: advised catch as output of the management procedure 

+ $B_{trigger}$: biomass trigger point of the HCR, specified as change point of biomass below which fishing mortality reduced relative to Ftgt. Btrigger is typically specified as ratio to $B_{tgt}$.

+ $B_{MSY}$: Biomass at $MSY$ as the "true" property on the Operating Model

+ $F_{MSY}$: Fishing mortality at $MSY$ as the "true" property on the Operating Model

# Short-cut MSE implementation in FLR  


Here, the MSE short-cut approach is implemented using the tools available in the Fisheries Library for R (FLR; Kell et al., 2007; https://flr-project.org/) 


## Set up `FLom` structure



Load OM conditioned to Stock Synthesis

```{r}
load("rdata/om.ss3ref.rdata",verbose=T)
```

Add 'true' MSY refpts 

```{r}


refpts= stka@refpts


```

```{r}  
stk=window(as(om.ref,"FLStock"))
```



Next set up the MSE horizon

```{r}
# FINAL year
fy <- 2045
# data year
dy <- dims(stk)$maxyear
# assessment year
ay = dy+1
# intermediate years
iy = ay:(ay+1)

```

For illustration the number of iterations are reduced to 500.

```{r}
# NUMBER iterations
it <- 500
```

Subset the 1000 simulated stock iterations to the first 100

```{r}
stk = iter(stk,1:it)
```

Get output from SS3 benchmark model for recruitment

```{r}
out=ss3out
yrs = an(dimnames(stka)$year)
recruit = out$recruit[out$recruit$Yr%in%yrs,]
sigmaR = out$sigma_R_in
dms <- list(year = yrs)
residuals <- FLQuant(exp(recruit$dev - 0.5 * recruit$biasadjuster *sigmaR^2), 
        dimnames = c(age = 0, dms), units = "")
recs = FLQuant(recruit$pred_recr, dimnames = c(age = 0, dms), units = "")


# Main recdevs
recyrs = recruit$Yr[recruit$era =="Main"]
maindevs = unitSums(residuals[,ac(recyrs)])
rho = cor(maindevs [,-1],maindevs [,-length(maindevs)])
s = round(an(bh@SV["s"]),3)

rho 
sigmaR
s

```

Get output from SS3 model specify assessment error on $SSB$ and $F$

```{r}
out = ss3out
# Short cut devs
ay = out$endyr # assessment year
SSBcv <- out$derived_quants$StdDev[out$derived_quants$Label==paste0("SSB_",ay)]/
  out$derived_quants$Value[out$derived_quants$Label==paste0("SSB_",ay)]

Fcv <- out$derived_quants$StdDev[out$derived_quants$Label==paste0("F_",ay)]/
  out$derived_quants$Value[out$derived_quants$Label==paste0("F_",ay)]

```

Summarize short-cut params

```{r}
spars = FLPar(s=s,sigmaR=sigmaR, rho=rho,Fcv=Fcv,SSBcv=SSBcv)

```

Generate recruitment 

```{r}
srdevs <- rlnormar1(n=it,sdlog=spars["sigmaR"], rho=spars["rho"], years=seq(dy, fy))

```

Now construct the `FLom` object from the `mse` package by passing on `FLStock`,`refpts`, `sr` and the method used for foward projections.

```{r}
om <- FLom(stock=stk, refpts=refpts, 
           sr=bh, projection=mseCtrl(method=fwd.om),deviances=srdevs)

class(om)
```

Next add the structure for the future years: average of last 3 years

```{r}
om <- fwdWindow(om, end=fy)

```

Make the last data year recruitment stochastic
```{r}
rec(stock(om))[, '2023'] <- rec(om)[1, '2023'] * srdevs[, '2023']
```

Project forward for intermediate year `iy` catch assumption of 8\% per year of the current 2022 catch with a 10\% implementation error.

```{r}
fiy= propagate(FLQuant(NA, dimnames=list(year=2024:2025)),it)

fiy[,ac(2024:2025)][] = fbar(stk)[,ac(2023)]*rlnorm(log(fbar(stk)[,ac(2023)]),0,0.1)

om  <- fwd(om,sr=bh,fbar = fiy)
```


```{r fig7, fig.height=9,fig.width=6,warning=FALSE,message=FALSE, fig.cap = paste0("OM trajectories with expanded MSE horizon, with black vertical line showing the assessment year and blue line the implementation year")}

plot(window(om,end=2025))+theme_bw()+
  geom_vline(xintercept = 2025,linetype=2,col=4)+
  geom_vline(xintercept = 2023,linetype=2)

```

Next, a so called observation error is constructed. In the case of the short-cut MSE, it simply holds the "perfect" stock information. For a full MSE with inbuilt estimation model  it would also generate the observations with errors, such a catch-at-age and survey numbers at age for SAM or a4a, or biomass surveys indices and catches for SPicT or JABBA.  

```{r}
oem <- FLoem(
  observations=list(stk=stock(om)),
  method=perfect.oem
)
```

However, there is increasing realisation that the assessment estimates are imperfect. Therefore, ICES has implemented procedures to add uncertainty about the key quantities $F$ and $SSB$, where the error on $F$ is specified by a the random error term $Fcv$ and a first order autocorrelation parameter $Fphi$ and the precision of $SSB$ can specified by $SSBcv$ 

Note, here the assessment errors are calculated from the OM, however, this could be also done from a retrospective analysis.


Short-cut deviations

```{r}
sdevs <- shortcut_devs(om, Fcv=Fcv, Fphi=0.432, SSBcv=SSBcv)
```

Finally, the implementation error module `iem` is setup. In this case, with a random catch implementation error of 10\%.

```{r}
iem <- FLiem(method=noise.iem,
  args=list(noise=rlnorm(it, rec(om) %=% 0, 0.1)))
```


```{r,eval=FALSE}
save(om, oem,sdevs,iem, file="rdata/flom.rda", compress="xz")
```

```{r,echo=FALSE}
load(file="rdata/flom.rda")
```

## Setting up harvest control rules

First the reference points are extracted are noted

```{r}
Ftgt = an(refpts$Fb35)
Btgt = an(refpts$Btgt)
Blim = an(refpts$Blim)
Bpa = an(refpts$Bpa)
MSY = an(refpts$Yeq)
Bmsy = an(refpts$Bmsy.om)
Fmsy = an(refpts$Fmsy.om)

```


In a first trial a fixed $F_{MSY}$ rule is applied to translate $F$-advice in to TACs, which is kept constant irrespective of biomass depletion.

```{r fig8, fig.height=6,fig.width=8,warning=FALSE,message=FALSE, fig.cap = paste0("Fixed $F_{adv} = F_{tgt}$ HCR, where $F_{tgt}=F_{MSY}$")}

pfix = plotGFCM(fadv = Ftgt,ftgt=Ftgt,btgt=Btgt,btrigger=0,bthr = Bpa,blim=Blim,kobe  =FALSE,text=F)+
  ylab("Fishing Mortality")+xlab("SSB")

pfix
```

This can be effectively implemented ICES advice rule hockey-stick by setting the $B_{trigger}$ to zero, using the `icesControl` function. This function can also take the `SSBdevs` and `Fdevs` that implement the deviations from the SSB and F with aim to account for assessment errors.



```{r}
arule <- mpCtrl(list(

  # (est)imation method: shortcut.sa + SSB deviances
  est = mseCtrl(method=shortcut.sa,
    args=list(SSBdevs=sdevs$SSB)),

  # hcr: hockeystick (fbar ~ ssb | lim, trigger, target, min)
  hcr = mseCtrl(method=hockeystick.hcr,
    args=list(lim=0, trigger=0.01, target=Ftgt,
    min=0, metric="ssb", output="fbar")),

  # (i)mplementation (sys)tem: tac.is (C ~ F) + F deviances
  isys = mseCtrl(method=tac.is,
    args=list(recyrs=-2, fmin=0, Fdevs=sdevs$F))
  ))


```

This rule can now be run passing on the `om`, `oem` and `arule` and an additional argument to set the implementation year to 2024.

Note that the default setting assumes 1 year lag between data (reference) year and assessment (reporting) year and that the TAC is implemented the next year. In case the assessment is conducted in 2023, based on data from 2022 and the TAC is implemented for 2024.



```{r,echo=FALSE}
load(file="rdata/om.ref.mp.sbr.500.rdata",verbose=T)
stk = stks$Fmsy.om

mseargs <- list(iy=2025, fy=fy, data_lag=1, management_lag=1, frq=1)

```

```{r,eval=FALSE,warning=FALSE}


system.time(
run <- mp(om, oem=oem, ctrl=arule, args=mseargs,verbose=F)
)

```


```{r fig9, fig.height=7,fig.width=9,warning=FALSE,message=FALSE, fig.cap = paste0("Initial OM and the MSE forecast horizon under a fixed $F_{tgt}$ rule")}

# make FLStocks from om until 2024 (implementation) and the run 
fmsy.stk = FLStocks(stock=window(om@stock,end=2025),
              fixedFmsy=stk)

plot(fmsy.stk )+facet_wrap(~qname,scales="free")+
  theme_bw()+
   geom_vline(xintercept = c(2023,2025),linetype=2,col=1)+
  geom_vline(xintercept = c(2030.5),linetype=2,col=2)+
 geom_vline(xintercept = c(2044.5),linetype=2,col=4)

```

The `run` can also be appended to the `om` to make a single `FLStockR` object with reference points

```{r}
runR = FLStockR(append(window(om@stock,end=2025),stk))
runR@refpts=refpts
```

This allows to quickly evaluate the stock status under a fixed $F_{MSY}$ rule.

It can be seen that despite "perfect" knowledge of the "true" $F_{MSY}$, and fishing pressure is on average $F_{MSY}$, the stock fails to attain biomass levels at $B_{MSY}$ with a relative high risk to fall below $B_{pa}$ and even $B_{lim}$. This is a well known fact as a result of the lags between data and management and assymetric risks in that exceeding $F_{MSY}$ is more consequential on both $SSB$ and long term yield, then fishing below $F_{MSY}$, In the case of the latter, more biomass is left in the water, which provides increased future reproduction potential and catch opportunity.   

```{r fig9a, fig.height=7,fig.width=9,warning=FALSE,message=FALSE, fig.cap = paste0("Stock Status under a fixed $F_{MSY}$")}

plotAdvice(runR)+
  geom_vline(xintercept = c(2023,2025),linetype=2)+
   geom_vline(xintercept = c(2030.5),linetype=2,col=2)+
 geom_vline(xintercept = c(2044.5),linetype=2,col=4)+
  scale_x_continuous(breaks=seq(1970,2045,5))+
  theme(axis.text.x = element_text(size=8, angle=90,vjust=0.5))
  

```

Even relatively simplified short-cut MSE frameworks provide a powerful to explore alternative HCRs to achieve better trade-off between risks and yield. 

Here, the conventional hockey-stick control rule is explored with different ratios of $F_{adv}$/$F_{tgt}$ and $B_{trigger}$/$B_{tgt}$ settings, where the $B_{trigger}$ promts a linear reduction in $F_{adv}$ if $SSB$ falls below it.



```{r , fig.height=6,fig.width=9,warning=FALSE,message=FALSE, fig.cap = paste0("Alternative HCR formulations, with $F_{min}=0$ (top) and $F_{min}=0.5F_{tgt}$ (bottom) under different $F_{adv}$ options set at $F_{adv}=F_{tgt}$, $F_{adv}=0.8F_{tgt}$ and at $F_{adv}=0.7F_{tgt}$ from left to right")}

fracFadv = rep(c(1,0.8,0.7),2) 
fracFmin= rep(c(0.0,0.5),each=3) 
Bclose = rep(c(0,Blim),each=3)

phcrs= Map(function(x,y,z){
p  = plotGFCM(fadv = Ftgt*y,btrigger =Btgt*0.5,ftgt=Ftgt,btgt=Btgt
         ,bthr = Bpa,blim=Blim,fmin=x*Ftgt,bclose=z,kobe  =FALSE,text=F)+
         ylab("Fishing Mortality")+xlab("SSB")
return(p)
},x=fracFmin,y=fracFadv,z=Bclose)
# plot ggplot list
ggarrange(plotlist=phcrs,nrow=2,ncol=3)

```

The same settings can be specified for the new `mps` function in `mse`, which allow to explore variations of the HCR parameters.

The function `combinations` enables to vary more than one parameter at the time. 

```{r}
hcrs =combinations(target=c(1,0.8,0.6)*Ftgt,
                   min=c(0.05,0.5)*Ftgt,trigger=0.5*Btgt)
hcrs$lim = rep(c(0,Blim),each=3)

# Add run with fixed Fmsy.om and Ftgt
hcrs$target = c(c(Fmsy,Ftgt),hcrs$target)
hcrs$trigger = c(c(0,0),hcrs$trigger)
hcrs$min = c(c(0,0),hcrs$min)
hcrs$lim = c(c(0,0),hcrs$lim)

hcrs
```

\pagebreak

## Implement OM with technical measures reducing catches below the MCRS of 33 cm.

The von Bertalanffy growth function:

```{r , message=FALSE,warning=FALSE,fig.height=4,fig.width=6, fig.cap = paste0("Von Bertalanffy function indicating the minimum size limit") }

linf = 62
k = 0.152 
t0 = -1.09
age = an(dimnames(stka)$age)
len = vonbert(linf,k,t0,age)
la = data.frame(age,len)

ggplot(la)+geom_line(aes(age,len))+theme_bw()+
  scale_x_continuous(breaks=age)+
  geom_hline(yintercept=33,linetype=2)+
  geom_vline(xintercept=c(3,4),linetype=2)+xlab("Age")+ylab("Length (cm)")
```

Adjust selectivity by reducing mortality of fish below the MCRS by 90\% 


```{r , message=FALSE,warning=FALSE,fig.height=4,fig.width=6, fig.cap = paste0("Comparison of current selectivity with a selectivity emulated by technical measures to reduce juvenile impact ") }

sel.cur =iterMedians(tail(yearMeans(catch.sel(om@stock),3)))
sel.tm = sel.cur
sel.tm[ac(0:3),] = sel.cur[ac(0:3),]*0.1

#compare
flqsel = FLQuants(sel.cur=sel.cur,sel.tm=sel.tm)

ggplot(flqsel)+geom_line(aes(age,data,col=qname,linetype=qname),linewidth=0.8)+theme_bw()+
  scale_x_continuous(breaks=age)+
  geom_vline(xintercept=c(3,4),linetype=2)+
  scale_color_manual(values = c(1,2))+theme(legend.title = element_blank())+
  xlab("Age")+ylab("Selectivity")
```

Implement technical measures in the OM from 2026 onwards 

```{r}
om.tm = om
om.tm@stock@harvest[,ac(2026:fy)][] = sel.tm/max(sel.tm)*Fmsy
```


Save MP input

```{r}
oms = list(ref=om,tm=om.tm)
mseargs <- list(iy=2025, fy=fy, data_lag=1, management_lag=1, frq=1)

save(oms,oem,arule,hcrs,mseargs,file="rdata/inpMP.sbr.500.rdata")
```


These changes in parameters can simply be passed on to the existing `arule` to run the variations with `mps`, e.g for the reference om  

```{r,eval=FALSE,warning=FALSE}
runs <- mps(om, oem=oem, ctrl=arule, args=mseargs,
  hcr=hcrs)

```

Note the this was run on a GFCM virtual environment in parallel

```{r,eval=F}


library(doParallel)
library(FLRef)
load("~/mseLite/sc_sbr/rdata/inpMP.sbr.500.rdata",verbose=T)
length(hcrs$target)
# Run in batches

ni = length(hcrs$target)
cl = ni*2

registerDoParallel(cl)
# indexing
o = rep(c(1,2),each=ni) # oms 
h = rep(1:ni,2)

start = Sys.time()

runs <- foreach(i = seq(ni*2)) %dopar% {
  
  
  # set stock index
  hcr = arule
  hcr$hcr@args$target = hcrs$target[h[i]]
  hcr$hcr@args$trigger = hcrs$trigger[h[i]]
  hcr$hcr@args$min = hcrs$min[h[i]]
  hcr$hcr@args$lim = hcrs$lim[h[i]]
  
  runi <- mp(oms[[o[i]]], oem=oem, ctrl=hcr , args=mseargs,verbose=T,parallel = FALSE)
  
  return(runi)
} # end of loop 
end = Sys.time()
time = end-start
time


scenarios = paste0(c(c("Fmsy.om","FixFtgt"),paste0(fracFadv,"Fadv.",fracFmin,"Fmin/Ftgt")),rep(c("",".tm"),each=ni))
names(runs) = scenarios

stks = FLStocks(lapply(runs,function(x){
  out = x@om@stock
  out = FLStockR(out)
  out@refpts = om@refpts
  out
}))

stkm = FLStocks(lapply(stks,function(x){
  stockMedians(x)
}))

plotAdvice(stkm)

save(stks ,file=paste0("rdata/om.refs.ftune.sbr.500.rdata"))

```



Now combine with the Fixed $F_{MSY}$ run and see if we can do better.

```{r}
scenarios = paste0(c(c("Fmsy.om","FixFtgt"),paste0(fracFadv,"Fadv.",fracFmin,"Fmin/Ftgt")),rep(c("",".tm"),each=length(hcrs$target)))
names(stks) = scenarios
```





# Performance Evaluation

## Evaluation horizons for preliminary rebuilding objectives

1. End of Management Plan 2030 (plus 1)

+ 70\% probability to be above $B_{lim}$ (high)


2. Intermediate 2031-2040

+ 90\% probability to be above $B_{lim}$ (very high)

+ 50\% probability to be above $B_{pa}$ (moderate)


3. Long-term rebuilding 2045 (two generation times)

+ 95\% probability to be above $B_{lim}$ (very high)

+ 80\% probability to be above $B_{pa}$ (high)

+ Mean biomass above 80\% $B_{tgt}$ 



## Performance statistics

+	Average Catch

+ Inter-annual variation in Catch Advice AAV(C)

+ Landing value as a function of Euro/kg at age (length)
    + Ages 0-3 = 15 Euro/kg (below 33 cm)
    + Ages 4-7 = 25 Euro/kg (33-45 cm)
    + Ages 7+ = 50 Euro/kg (larger 45 cm)


+	Probability above Blim  
  +	2031: Red line set to 0.7 (i.e. 70\% $B{lim}>B$)
  +	2031-2040: Red line set to 0.95 (i.e. 90\% $B{lim}>B$) 
  +	2045: Red line set to 0.95 (i.e. 95\% $B{lim}>B$) 
  
+	Probability above Bpa 
  +	2031-2040: Red line set to 0.5 (i.e. 50\% $B{pa}>B$)
  +	2045: Red line set to 0.8 (i.e. 80\% $B{pa}>B$)

+	Mean $B/B_{tgt}$ 
  +	2045: Red line set to 0.8 (i.e $B/B{tgt}>0.8$ 
  +	2045: Green line set to $B/B{tgt}>1$ 
  

  
Landing value function  
  
```{r , message=FALSE,warning=FALSE,fig.height=4,fig.width=6, fig.cap = paste0("Von Bertalanffy function indicating the assumed breaks in marked price categories (dashed lines)") }

ggplot(la)+geom_line(aes(age,len))+theme_bw()+
  scale_x_continuous(breaks=age)+
  geom_hline(yintercept=c(33,45),linetype=2)+
  geom_vline(xintercept=c(4,7.5),linetype=2,col=2)+
  xlab("Age")+ylab("Length (cm)")
```  
  
```{r}
sbr.value = function(stock){
lval = yearMeans(mat(stock))
lval[] = 50
lval[ac(0:3),] = 15
lval[ac(4:7),] = 25
lvq = catch.n(stock)
lvq[] = lval
return(quantSums(catch.n(stock)*catch.wt(stock)*lvq))
  }
```  
  
```{r}
metrics <- list(SB = function(x)unitSums(ssb(x)), 
                F = function(x)unitMeans(fbar(x)), 
                C = function(x)unitSums(catch(x)),
                Rec= function(x)unitSums(rec(x)),
                Val = function(x)unitSums(sbr.value(x)))
```  
  
  

```{r}
stats <- list(

  # mean(C)
  A.C=list(~yearMeans(C), name="mean(Catch)",
    desc="1 Mean catch over years"),

  B.Lval = list(~yearMeans(Val), name="Landing[Euro]",
    desc="Mean catch over years"),
 
  # AVVC
  C.AAVC=list(~yearMeans(abs(C[, -1] - C[, -dim(C)[2]])/C[, -1]),
    name="AAV(Catch)", desc="Average annual variability in catch"),

  
  # P(SB < SBlim) at least once
  D.PBlim=list(~yearMeans(iterMeans(((SB/Blim) > 1))),
    name="SB>B[limit]",
    desc="Probability that spawner biomass is above Blim"),
 
  
  # P(SB>Bpa)
  E.PBpa=list(~yearMeans(iterMeans((SB/Bpa) > 1)), name="P(SB>SB[pa])",
    desc=" Probability that spawner biomass is above Bpa"),

  
  
  # E.PBlim= list(~apply(iterMeans((SB/Blim) > 1),1,min), 
  #  name="P3(B<Blim)", desc="Probability that SSB > Blim"),
  
  F.BBmsy=list(~yearMeans((SB/Btgt)), name="B/B[tgt]",
    desc="B/Btgt"),
  
  G.PFmsy=list(~yearMeans(iterMeans((F/Fmsy)<1)), name="P(F<F[MSY])",
    desc="F<Fmsy")

  
    
)

```


\pagebreak

## Performance evaluation of OM reference runs



```{r , fig.height=7,fig.width=9,warning=FALSE,message=FALSE, fig.cap = paste0("Refernec OM showing the assessment and MSE forecast horizon under a fixed $F_{MSY}$ $F_{tgt}$ rule and 6 HCR hockey-stick scenarios")}

ref = FLStockR(window(stock(om),end=2025))
ref@refpts = refpts


pstks = FLStocks(c(FLStocks(stock=window(ref,end=2025)),
              stks[1:8]))

plotAdvice(pstks)+facet_wrap(~qname,scales="free")+
  theme_bw()+
  scale_color_manual(values=c("black",ss3col(length(pstks))[-1]))+
  scale_fill_manual(values=c("darkgrey",ss3col(length(pstks))[-1]))+
  geom_vline(xintercept = 2023,linetype=2,col=1)+
  geom_vline(xintercept = 2030.5,linetype=2,col=2)+
  geom_vline(xintercept = 2044.5,linetype=2,col=4)+
  scale_x_continuous(breaks=seq(1970,2045,5))+
  theme(axis.text.x = element_text(size=8, angle=90,vjust=0.5))

```

```{r, fig.height=7,fig.width=9,warning=FALSE,message=FALSE, fig.cap = paste0("Median MSE forecast horizon under a fixed $F_{tgt}$ rule and 6 HCR scenarios")}

medstks = FLStocks(lapply(stks,function(x){
  out = stockMedians(x)
  out@refpts = x@refpts[-c(4:5)] # rmove R0 and B0 
  out
  }))
medref = stockMedians(ref)
# for plotting with assessment
pmstks = FLStocks(c(FLStocks(stock=window(medref,end=2025)),
              medstks))

plotAdvice(medstks[1:8] )+facet_wrap(~qname,scales="free")+
  scale_color_manual(values=c(ss3col(length(medstks[1:8]))))+
  theme_bw()+xlab("Year")+
  geom_vline(xintercept = 2023,linetype=2,col=1)+
  geom_vline(xintercept = 2030.5,linetype=2,col=2)+
  geom_vline(xintercept = 2044.5,linetype=2,col=4)+
  scale_x_continuous(breaks=seq(1970,2045,5))+
  theme(axis.text.x = element_text(size=8, angle=90,vjust=0.5))

```


\pagebreak 


### End of long-term management plan 2030 (plus 1)

Compute performance metrics

```{r}
perf_map <- performance(stks[(1:8)], statistics=stats[c(1,4)],metrics=metrics, years=list(2031))

```

PLOT performance


```{r, fig.height=4,fig.width=8,warning=FALSE,message=FALSE, fig.cap = paste0("Performance under the OM reference run: Long-term through 2045")}

ncol = length(unique(perf_map$mp)) 
plotBPs(perf_map,
        limit=c(D.PBlim=0.7))+
  facet_wrap(~name,scale="free_y",ncol=2)+
  ylim(c(0, NA))+theme_bw()+
  ggtitle(paste0("Performance: End of MAP 2030"))+
  ylab("Performance statistics")+
  scale_fill_manual(values=ss3col(ncol))+ # USE FLRef::ss3col
 theme(axis.text.x=element_blank())+xlab("Candidates")

```

\pagebreak


### Intermediate 2031-2040


Compute performance metrics

```{r}
perf_int <- performance(stks[c(1:8)], statistics=stats[c(1,2,3,4,5,7)],metrics=metrics, years=list(2031:2040))

```

PLOT performance

```{r, fig.height=7,fig.width=8,warning=FALSE,message=FALSE, fig.cap = paste0("Performance under the OM reference run: Itermediate 2031-2040")}

ncol = length(unique(perf_int$mp)) 
plotBPs(perf_int,
        limit=c(D.PBlim=0.9,E.PBpa=0.5,G.PFmsy=0.8),target = c(G.PFmsy=1))+
  facet_wrap(~name,scale="free_y",ncol=2)+
  ylim(c(0, NA))+theme_bw()+
  ggtitle(paste0("Performance: Intermediate 2031-2040"))+
  ylab("Performance statistics")+
  scale_fill_manual(values=ss3col(ncol))+ # USE FLRef::ss3col
 theme(axis.text.x=element_blank())+xlab("Candidates")

```



```{r, fig.height=6,fig.width=7,warning=FALSE,message=FALSE, fig.cap = paste0("Trade-offs under the OM reference run betweem betweem mean landing value (in 1000 Euro) and risk of $SSB$ falling below $Blim$")}

fmsy.val = median(perf_int[perf_int$run=="Fmsy.om" &perf_int$statistic== "B.Lval",]$data)

plotTOs(perf_int, x="B.Lval", y=c("D.PBlim"))+
  geom_hline(yintercept = 0.9,col=2,linetype=2)+
  geom_vline(xintercept = fmsy.val,col=4,linetype=2)+
  theme_bw()+xlab("Landing Value in 1000 Euro")+
  ylab("P(B>Blim)")+
    scale_fill_manual(values=ss3col(ncol))+scale_y_continuous(limits = c(0,1))
```




```{r,echo=F,message=F,warning=F}
dir.create("perftabs",showWarnings = F)
pf = perf_int
hcrs.ref = c(1:8)
df.pf = aggregate(data~mp+name,pf,mean)
sd.pf = aggregate(data~mp+name,pf,sd)
sd.pf = sd.pf[sd.pf$name=="mean(Catch)",] 
tab.pf = cbind(reshape2::dcast(df.pf,mp~name),SD.Catch=sd.pf$data)
# Add Fadv and Btrigger
tab.pf = cbind(tab.pf,Fadv=hcrs$target[hcrs.ref],
               Btrigger=hcrs$trigger[hcrs.ref],
                    Fmin=hcrs$min[hcrs.ref])


tab.pf = tab.pf[,c(1,9:11,3,4,8,2,7,6)]

tab.pf[,-1] = round(tab.pf[,-1],3)
tab.pf[c(3,5:7)] = round(tab.pf[c(3,5:7)],1)
tab.pf$Fmin=  ifelse(tab.pf$Btrigger==0,tab.pf$Fadv,tab.pf$Fmin)
tab.pf$mp = c(rep("FixedF",2),rep("HCR.F0",3),rep("HCR.Fmin",3))
names(tab.pf)

names(tab.pf) = c("MP","Fadv","Btri","Fmin","Value(Euro)","Catch","SD","AVV","B>Blim","B>Bpa")

# Save csv
write.csv(tab.pf,file="perftabs/perf.int.ref.tab.csv",row.names = F)



knitr::kable(tab.pf,"pipe",
      align ="lccccc",
      caption="Comparison of MP performance under the OM reference run: Itermediate 2031-2040")


```

\pagebreak

### Long-term biomass rebuilding 2045 (2 Generation times)


Compute performance metrics

```{r}
perf_end <- performance(stks[-c(1:8)], statistics=stats[c(4,5,6)],metrics=metrics, years=list(2045))

```

PLOT performance


```{r, fig.height=4,fig.width=10,warning=FALSE,message=FALSE, fig.cap = paste0("Performance under OM reference runs: Long-term through 2045")}

ncol = length(unique(perf_end$mp)) 
plotBPs(perf_end,
        limit=c(D.PBlim=0.95,E.PBpa=0.8,F.BBmsy=0.8),target = c(F.BBmsy=1))+
  facet_wrap(~name,scale="free_y",ncol=3)+
  ylim(c(0, NA))+theme_bw()+
  ggtitle(paste0("Performance: Long-term 2045 (2xGT)"))+
  ylab("Performance statistics")+
  scale_fill_manual(values=ss3col(ncol))+ # USE FLRef::ss3col
 theme(axis.text.x=element_blank())+xlab("Candidates")

```

\pagebreak

## Performance under additional technical measures to reduce catches below the MCRS

Emulated techical measures (tm) effect on selectivity

```{r , message=FALSE,warning=FALSE,fig.height=4,fig.width=6, fig.cap = paste0("Comparison of current selectivity with a selectivity emulated by technical measures to reduce juvenile impact ") }

ggplot(flqsel)+geom_line(aes(age,data,col=qname,linetype=qname),linewidth=0.8)+theme_bw()+
  scale_x_continuous(breaks=age)+
  geom_vline(xintercept=c(3,4),linetype=2)+
  scale_color_manual(values = c(1,2))+theme(legend.title = element_blank())+
  xlab("Age")+ylab("Selectivity")
```

Select the HCRs with $F_{adv} = F_{tgt}$ for comparison

```{r , fig.height=3,fig.width=9,warning=FALSE,message=FALSE, fig.cap = paste0("Selected HCRs for comparison to scenarios with additional technical measure (tm)")}
ggarrange(pfix,phcrs[[1]],phcrs[[4]],nrow=1,ncol=3)+
  ylab("Fishing Mortality")+xlab("SSB")


```
Select HCRs runs 

```{r}
#names(stks)
tmsel =c(2,10,3,11,6,14)
stks.tm = stks[tmsel] 
```



```{r , fig.height=7,fig.width=9,warning=FALSE,message=FALSE, fig.cap = paste0("Comparison of the reference OM with additional technical measures (tm) implementation showing the median MSE forecast horizon under a fixed $F_{tgt}$ rule and two selected HCR scenarios")}

medstks.tm = FLStocks(lapply(stks.tm,function(x){
  out = stockMedians(x)
  out@refpts = x@refpts[-c(4:5)] # rmove R0 and B0 
  out
  }))


plotAdvice(medstks.tm )+facet_wrap(~qname,scales="free")+
  scale_color_manual(values=c(ss3col(length(medstks.tm))))+
  theme_bw()+xlab("Year")+
  geom_vline(xintercept = 2023,linetype=2,col=1)+
  geom_vline(xintercept = 2030.5,linetype=2,col=2)+
  geom_vline(xintercept = 2044.5,linetype=2,col=4)+
  scale_x_continuous(breaks=seq(1970,2045,5))+
  theme(axis.text.x = element_text(size=8, angle=90,vjust=0.5))

```


\pagebreak 


### End of long-term management plan 2030 (plus) with technical measures

Compute performance metrics

```{r}
perf_map.tm <- performance(stks.tm, statistics=stats[c(1,4)],metrics=metrics, years=list(2031))

```

PLOT performance


```{r, fig.height=4,fig.width=8,warning=FALSE,message=FALSE, fig.cap = paste0("Comparison of performance under the OM reference run with additional technical measures (tm): Long-term through 2045")}

ncol = length(unique(perf_map.tm$mp)) 
plotBPs(perf_map.tm,
        limit=c(D.PBlim=0.7))+
  facet_wrap(~name,scale="free_y",ncol=2)+
  ylim(c(0, NA))+theme_bw()+
  ggtitle(paste0("Performance: End of MAP 2030"))+
  ylab("Performance statistics")+
  scale_fill_manual(values=ss3col(ncol))+ # USE FLRef::ss3col
 theme(axis.text.x=element_blank())+xlab("Candidates")

```


\pagebreak


### Intermediate 2031-2040


Compute performance metrics

```{r}
perf_int.tm <- performance(stks.tm, statistics=stats[c(1,2,3,4,5,7)],metrics=metrics, years=list(2031:2040))

```

PLOT performance

```{r, fig.height=7,fig.width=8,warning=FALSE,message=FALSE, fig.cap = paste0("Comparison of performance under the OM reference run with additional technical measures (tm): Itermediate 2031-2040")}

ncol = length(unique(perf_int.tm$mp)) 
plotBPs(perf_int.tm,
        limit=c(D.PBlim=0.9,E.PBpa=0.5,G.PFmsy=0.8),target = c(G.PFmsy=1))+
  facet_wrap(~name,scale="free_y",ncol=2)+
  ylim(c(0, NA))+theme_bw()+
  ggtitle(paste0("Performance: Intermediate 2031-2040"))+
  ylab("Performance statistics")+
  scale_fill_manual(values=ss3col(ncol))+ # USE FLRef::ss3col
 theme(axis.text.x=element_blank())+xlab("Candidates")

```



```{r, fig.height=6,fig.width=7,warning=FALSE,message=FALSE, fig.cap = paste0("Trade-offs under the OM reference run and additional technical measures (tm) betweem mean landing value (in 1000 Euro) and risk of $SSB$ falling below $Blim$")}

fmsy.val = median(perf_int[perf_int$run=="Fmsy.om" &perf_int$statistic== "B.Lval",]$data)

plotTOs(perf_int.tm, x="B.Lval", y=c("D.PBlim"))+
  geom_hline(yintercept = 0.9,col=2,linetype=2)+
  geom_vline(xintercept = fmsy.val,col=4,linetype=2)+
  theme_bw()+xlab("Landing Value in 1000 Euro")+
  ylab("P(B>Blim)")+
    scale_fill_manual(values=ss3col(ncol))+scale_y_continuous(limits = c(0,1))
```



Generate performance Table



```{r,echo=F,message=F,warning=F}
dir.create("perftabs",showWarnings = F)
pf = perf_int.tm 
hcrs.tm = c(2,3,6)
df.pf = aggregate(data~mp+name,pf,mean)
sd.pf = aggregate(data~mp+name,pf,sd)
sd.pf = sd.pf[sd.pf$name=="mean(Catch)",] 
tab.pf = cbind(reshape2::dcast(df.pf,mp~name),SD.Catch=sd.pf$data)
# Add Fadv and Btrigger
tab.pf = cbind(tab.pf,Fadv=hcrs$target[hcrs.tm],
               Btrigger=hcrs$trigger[hcrs.tm],
                    Fmin=hcrs$min[hcrs.tm])
names(tab.pf)

tab.pf = tab.pf[,c(1,9:11,3,4,8,2,7,6)]

tab.pf[,-1] = round(tab.pf[,-1],3)
tab.pf[c(3,5:7)] = round(tab.pf[c(3,5:7)],1)
tab.pf$Fmin=  ifelse(tab.pf$Btrigger==0,tab.pf$Fadv,tab.pf$Fmin)
tab.pf$mp = rep(c("Ref","TM"),3)
names(tab.pf) = c("MP","Fadv","Btri","Fmin","Value(Euro)","Catch","SD","AVV","B>Blim","B>Bpa")
# Save csv
write.csv(tab.pf,file="perftabs/perf.int.tm.tab.csv",row.names = F)



knitr::kable(tab.pf,"pipe",
      align ="lccccc",
      caption="Comparison of performance under the OM reference run with additional technical measures (tm): Itermediate 2031-2040")


```


\pagebreak

### Long-term biomass rebuilding 2045 (2 Generation times)


Compute performance metrics

```{r}
perf_end.tm <- performance(stks.tm, statistics=stats[c(4,5,6)],metrics=metrics, years=list(2045))

```

PLOT performance


```{r, fig.height=4,fig.width=10,warning=FALSE,message=FALSE, fig.cap = paste0("Comparison of performance under the OM reference run with additional technical measures (tm): Long-term through 2045")}

ncol = length(unique(perf_end.tm$mp)) 
plotBPs(perf_end.tm,
        limit=c(D.PBlim=0.95,E.PBpa=0.8,F.BBmsy=0.8),target = c(F.BBmsy=1))+
  facet_wrap(~name,scale="free_y",ncol=3)+
  ylim(c(0, NA))+theme_bw()+
  ggtitle(paste0("Performance: Long-term 2045 (2xGT)"))+
  ylab("Performance statistics")+
  scale_fill_manual(values=ss3col(ncol))+ # USE FLRef::ss3col
 theme(axis.text.x=element_blank())+xlab("Candidates")

```

\pagebreak




# Robustness Tests


## Impaired recruitment scenarios

The success for rebuilding stocks from low biomass levels strongly depend future recruitment. Therefore, the assumptions about future recruitment are perhaps the most critical for robustness testing of rebuilding management strategies.    

### low steepness $s = 0.6$

```{r}
flsr = as.FLSR(stka,model=bevholt)
spr0 = mean(spr0y(stka))
R0  = an(bh@SV["R0"])
bh.low = srrTMB(flsr,spr0=spr0,s=0.6,s.est = FALSE,r0=c(R0,0.000001))

# parameter
params(bh.low)
bh.low@SV
```
```{r}

refpts = stka@refpts
rpts.ref = computeFbrp(stka,bh,proxy=c("msy","bx"),x=35,type="btgt",bpa=0.5*Blim,blim=Blim)
rpts.s06 = computeFbrp(stka,bh.low,proxy=c("msy","bx"),x=35,type="btgt",bpa=0.5*Blim,blim=Blim)
refpts.s06 = refpts
refpts.s06[c("Fmsy.om","Bmsy.om")][] = Fbrp(rpts.s06)[1:2]
```

```{r ,fig.width=8, fig.height=7, message=FALSE, fig.cap = paste0("Comparison of feference point estimates between the reference Om and the low steepness OM shown as functions of $SSB$, $F$, Yield and Recruitment. Grey dots show the corresponding observations for the annual SS3 reference model") }

ploteq(FLBRPs(ref=rpts.ref ,"low.s=0.6"=rpts.s06))

```
\pagebreak

### Depensation


```{r}
flsrd = as.FLSR(stka,model=bevholtDa)
bh.dep = srrTMB(flsrd,spr0=spr0,s=s,s.est = T,r0=c(R0,0.1),d=1.5,d.est=F)


```

### Compare recruitment scenarios

```{r , message=FALSE,warning=FALSE,fig.height=4,fig.width=6, fig.cap = paste0("Comparison of 3 different stock-recruitment assumptions") }
plotsrs(FLSRs(ref=bh,"low.s=0.6"=bh.low,depensation=bh.dep),b0=T)+
 geom_hline(yintercept = R0,linetype=2)+
  geom_vline(xintercept = an(om@refpts["B0"]),linetype=2)

```



Make deterministic forecast for mean expectation from the recruitment scenarios under a fixed $F_{tgt}$ scenario

```{r}
srs= FLSRs(reference=bh,low.s=bh.low,depensation=bh.dep)
fstk = stf(stka,22)

fws <- FLStocks(lapply(srs,function(x){
  
 res =  FLStockR(fwd(fstk, sr=x,
            control=fwdControl(year=2023:2045,
            value=Ftgt,quant="f")))
 res@refpts = stka@refpts[1:7]
 return(res)
}))


```


```{r , message=FALSE,warning=FALSE,fig.height=6,fig.width=8, fig.cap = paste0("Comparison of deterministic forecasts under 3 different recruitment scenarios with a fixed $F_{tgt}$") }

plotAdvice(fws)+
  scale_x_continuous(breaks=seq(0,3000,5))+
  theme(axis.text.x = element_text(size=8, angle=90,vjust=0.5))

```

\pagebreak

## TODO: Implement robustness MSE runs


```{r}
om.s06 = om
om.dep = om
om.s06@sr = bh.low
om.dep@sr = bh.dep

oms.rob = list(om.s06=om.s06,om.dep=om.dep)
mseargs <- list(iy=dy, fy=fy, data_lag=1, management_lag=1, frq=1)

save(oms.rob,oem,arule,hcrs,mseargs,file="rdata/inpMP.sbr.rob.500.rdata")
```

